---
title: "Almond yield as a function"
author: "Halina Do-Linh"
date: "4/10/2022"
output: html_document
---

```{r setup, include=FALSE}
# set up
knitr::opts_chunk$set(echo = TRUE)

# attached packages
librarian::shelf(
  tidyverse,
  here,
  janitor,
  patchwork
)

# read in data txt file
# had to add "d" in clim.txt as dummy column to get data lined up correctly
# wy = water year (oct 1 = sept 30)
# wyd = water year day
clim_data <- read_table(here("hw/data/clim.txt")) %>% 
  clean_names() %>% 
  select(-d) # dropped "d" aka index column
```

# Almond yield as a function

## Assignment

For assignment two we will be building the almond model (from Lobell et al. 2006) in R. As a side note: there are always multiple ways to code something in R; of course focus on getting the correct answer first, but also remember that we want to strive for our code being as simple and streamline as possible. Style counts. Make sure you choose meaningful variable names and add comments. Include comments at the top of the function to tell the user what the inputs/outputs are and their units and format.

Here are some ideas to think though. First, the climate data is going to need to be subsetted. How the almond function is written will dictate how the climate data is going to be stored, after it is subsetted. Second, we want to build a clean function that is versatile. Here are two model outlines to follow

- Almond_model <- function(clim_var1, clim_var2, parameters){……}

- Almond_model <- function(clim, parameters){……}

The first example is where the climate variables are separately input into the function, and the second is where a data frame is the input in the function and you extract the useful data from it. The first demands that the data is subset beforehand, the second subsets the data as part of the function, but demands the dataframe be structured in a specific way. There are advantages to both – the first is simpler model to ‘code’; but requires more ‘beforehand’ work by the user. You can pick which option you prefer (or try both).

# Steps:

1. Code your function in R; save it as a separate file called “the name of your function”.R; Make sure you include documentation Store your R function in a git repository - you will need it again for future assignment 

2. In an Rmarkdown document, write code to read in the “clim.txt”, (which you can find on [esm232_examples github site]{https://github.com/naomitague/ESM232_Examples.git}. Do any subsetting required for input to your function, run the function and save results.

3. Summarize the results as follows

The end product we are looking for: calculate the almond yield anomaly for each year, and summarize the data. We will leave how to summarize the data up to you. Think about what metrics would be best to explain the trends in outputs. Tables, graphs, and summary metrics are all acceptable – these you will generate outside the function in your Rmarkdown. Make sure to include axis labels, units and captions.

4. Finally, write one paragraph summarizing your findings. Submit pdf of rmarkdown.and the function.R file to gauchospace – Submit as a group

Three answers to check your model against:

- 2000: 9.59
- 2001: 159.51
- 2002: 0.24

# My Function

Questions:
- did anyone feel like the data was read in wrong?
- tmax_c is year?

```{r}
# need feb and jan data
almond_yield <- function(temp, precip,
                         temp_min_coeff_1 = -0.015, 
                         temp_min_coeff_2 = -0.0046, 
                         percip_coeff_1 = -0.07, 
                         percip_coeff_2 = 0.0043, 
                         intercept = 0.28){
  
  # calculating almond yield
  yield = (temp_min_coeff_1 * temp) + (temp_min_coeff_2 * temp^2) +
    (percip_coeff_1 * precip) + (percip_coeff_2 * precip^2) + intercept
  
  # results
  return(yield)
  
  } # EO function
```


```{r}
# necessary climate variables
temp_feb <- clim_data %>% 
  group_by(month, year) %>% 
  summarize(mean_tmin_c = mean(tmin_c)) %>% 
  filter(month == 2)
  
precip_jan <- clim_data %>% 
  group_by(month, year) %>% 
  summarize(mean_precip = sum(precip)) %>% 
  filter(month == 1) 
```


```{r}
## ANSWERS ##
# 2000: 9.59
# 2001: 159.51
# 2002: 0.24

yield_2000 <- almond_yield(temp = temp_feb$mean_tmin_c[[12]], 
                           precip = precip_jan$mean_precip[[12]])

yield_2001 <- almond_yield(temp = temp_feb$mean_tmin_c[[13]], 
                           precip = precip_jan$mean_precip[[13]])

yield_2002 <- almond_yield(temp = temp_feb$mean_tmin_c[[14]], 
                           precip = precip_jan$mean_precip[[14]])

yield_2000
yield_2001
yield_2002
```


```{r}
# join feb and jan data frames
joined_tempfeb_precipjan <- left_join(precip_jan, temp_feb, by = "year")

# cleaned joined df
joined_clean <- joined_tempfeb_precipjan %>% 
  select(-c(month.x, month.y)) %>% 
  rename(mean_precip_jan = mean_precip,
         mean_tmin_c_feb = mean_tmin_c)
```


```{r}
joined_clean_almond_yield <- joined_clean %>% 
  mutate(almond_yield_pct = almond_yield(temp = mean_tmin_c_feb,
                                        precip = mean_precip_jan))

almond_plot <- ggplot() +
  geom_line(data = joined_clean_almond_yield, aes(x = year, y = almond_yield_pct),
            color = "brown") +
  labs(title = "",
       x = "Year",
       y = "% Anomaly") +
  theme_minimal() 

precip_plot <- ggplot() +
  geom_line(data = joined_clean_almond_yield, aes(x = year, y = mean_precip_jan),
            color = "cyan4") +
  labs(x = "Year",
       y = "Precip (mm)") +
  theme_minimal()

temp_plot <- ggplot() +
  geom_line(data = joined_clean_almond_yield, aes(x = year, y = mean_tmin_c_feb), 
            color = "red") +
  labs(x = "Year",
       y = "Temp (\u00B0C)") +
  theme_minimal()
  
almond_plot / precip_plot / temp_plot
```

# Summary

We found the annual percent anomaly yield for almonds based on the model provided by Lobell et al. 2006 for the years 1988-2010. Based on the function provided by the paper, we see that the year 1995 had a result of ~2000% increase in percent anomaly yield for almonds. Through the time period we see other slight spikes in percent anomaly yield of ~250% - 600% for the years 1997, 2005, and 2008. We then added the precipitation data for the same years, and saw that there were spikes that correspond to the spikes in almond anomalies. The temperature anomalies did not line up as precisely as the precipitation anomalies did. There are similar spikes in the years that had larger almond yields, however there are some spikes in temperature that did not line up with increased almond yield. For example, the year 2000 had a high temperature, but did not have an increase in almond production. When looking at these three plots we can assume that precipitation has a large effect on almond yield compared to temperature. Although we can not rule out that temperature might have an effect on precipitation.   
